---
title: 'Exercise 1: R basics'
author: "Fursham Hamid"
date: "`r Sys.Date()`"
output: 
  html_document:
            number_sections: FALSE
            theme: paper
header-includes:
            \usepackage{graphicx,amsmath,caption,multicol}
fontsize:
        12pt 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Exercise instructions

Start RStudio and follow along with the instructions for each question. Try each question
out on your own and check your answers with the answer given below each question.

## Arithmetic operations and R  variables

### 1. 
Perform the following calculation in R: 
$\frac{25}{75+25}$  
What value did you obtain?

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
25/(75+25)
```
Answer is 0.25
</details> 

### 2. 
Repeat the above calculation, but save the returned value
into a variable called `x`.

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
x <- 25/(75+25)
```
</details> 

### 3.
What is the data type of variable `x`? How do you
check for this information? Name 3 other data structures
in R.

<details>
  <summary>Click for Answer!</summary>
  `x` is a numerical variable. There are several ways
  to check for this. First, you may run the function
  `typeof()` as shown below.
```{r eval=F}
typeof(x)
```
   Secondly, you may check RStudio's "Environment" panel
   for the data type under column "Type".  
   <br>
   Other common R data structures include:
   
   1. Character variables
   2. Logical variables
   4. Atomic vectors.
   5. Lists.
   6. Factors
   7. Matrices
   8. DataFrames.
   9. R Objects
</details> 

### 4.
Consider the following code:
```{r eval=F}
x*x
```

Write an alternative code to perform the above operation.
What is the value of `x` now? Why is this so?

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
x^2
```
`x` still holds the value of 0.25. This is because the code above
does not replace the original value of `x`.
</details> 

## Installing and loading packages

The [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org/) 
repository homes a ton of stable R packages as well as sample datasets. 

### 5.
Install `nycflights13` package on your computer.

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
install.packages("nycflights13")
```
</details> 

### 6.
Show one way to check if `nycflights13` package is
successfully installed.

<details>
  <summary>Click for Answer!</summary>
  One way to test for proper installation is to load
  the package in your environment. If no error appear,
  then the pacakge is successfully installed.
```{r eval=F}
library("nycflights13")
```
</details> 

## DataFrame and Vectors

### 7.
`flights` is one of the dataset found in `nycflights13` 
package. Using code, determine its data type
and dimension. 

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
# checking data type
typeof(flights)

# checking dimension
dim(flights)
```
</details> 


Run the following code to extract the column `dep_time` 
from the `flights` dataset (from nycflights13 pacakge).

```{r eval=F}
v <- flights$dep_time
```


### 8.
How do you add 100 to each element of vector `v`?

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
v <- v + 100
```
The above code works because R employ performs
operations using vectorisation.

</details> 

### 9.
How do you extract the first 5 elements of vector `v`?

<details>
  <summary>Click for Answer!</summary>
```{r eval=F}
v[1:5]
```
</details> 


### 10.
What does the code below do? What is the output?
```{r eval=F}
flights[100,10]
```

<details>
  <summary>Click for Answer!</summary>
  The code above wil extract the value at the 100th row and the 10th column.  
  It outputs a dataframe with1 single value "US".
</details> 


<br/>
__End of exercise__









































